Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-02-22 resex.tex
fixed merge errors for bibliography
Revised section III A on residual kicks.

Currently, no HEL is installed in the LHC, and the kick on the proton beam core must therefore be emulated with the help of other devices already installed in the LHC. For this purpose the kick is approximated to first order by a dipole kick. During the experiments presented in this paper, the dipole kick with the corresponding excitation pattern could then be applied with the LHC transverse damper as will be described in detail in Section~\ref{sec:adt}. As will be shown in Sec.~\ref{core:sec:1} and \ref{core:sec:2}, the expected kick to first order from the HEL bends is estimated to be~(Eq.~\ref{eqn:kick_bends}):
\begin{equation}
	\Delta x'_{\rm{bends}},\Delta y'_{\rm{bends}}\approx \q{0.5}{nrad},
\end{equation}
and for the central region due to profile imperfections~(Eq.~\ref{eqn:kick_central}):
\begin{equation}
	\Delta x'_{\rm{central \ region}}, \Delta y'_{\rm{central \ region}} \approx \q{20}{nrad} % this should be rounded to the same number of digits as the 0.5nrad
\end{equation}
where the current HEL design parameters are used (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}). The contribution from the central region is clearly dominating.

\subsection{Uncompensated kicks from HEL bends}
\label{core:sec:1}
As reference for the estimate of the dipole component originating from the HEL bends, the symplectic map derived in \cite{hel_bends_stancari} is used. In this case the HEL bends are modeled as a bent cylindrical pipe with a static charge distribution of 1~A and 5~keV electrons deflected in the horizontal plane, where the parameters were her still based on the old gun design for the Tevatron electron lens \textcolor{red}{Giulio, is this correct?}. In case of an U-shaped HEL, the transverse dipole kicks at entrance and exit add up, while for a S-shape the kicks compensate each other. For this reason a S-shape has been chosen for the HL-LHC HEL (see Fig.~\ref{fig:hel_layout}). Assuming and S-shaped HEL, uncompensated kicks therefore only arise from imperfections, which can originate from profile imperfections and current fluctuations. As a first estimate, we assume in this paper 10\% fluctuations between the entrance and exit and, in addition, that the kicks from entrance and exit due to these imperfections add up. Using the electric field calculations in \cite{hel_bends_stancari}, the maximum values for the integrated electric field are then
\begin{equation}
\int_{z_1}^{z_2} E_{x,y} dz= 10 \ \mathrm{kV}.
\end{equation}
%given an \q{1}{A}, \q{5}{keV} electron beam.
%For \q{7}{TeV} protons and neglecting magnetic effects, this yields a kick of:
%\begin{equation}
%\Delta x'=\Delta y'= \q{1.4}{nrad},
%\end{equation}
%Scaling to the HEL design parameters of a \q{5}{A} and \q{10}{keV} electron beam (Table~\ref{tab:hel_param}), one obtains:
%\begin{equation}
%\int_{z_1}^{z_2} E_{x} dz= 36 \ \mathrm{kV} \Rightarrow \Delta x'= \q{5.1}{nrad}
%\end{equation}
Neglecting magnetic effects and scaling to HL-LHC and HEL design parameters (Tables~\ref{tab:hllhc_param}--\ref{tab:hel_param}) yields a kick of \cite{md_sim_hel_res_ex_fitterer}:
\begin{equation}
\int_{z_1}^{z_2} E_{y} dz= 36 \ \mathrm{kV} \Rightarrow \Delta y'= \q{5.1}{nrad}
\end{equation}
Assuming now 10\% fluctuation between entrance and exit kick, the expected kick from the HEL bends is
\begin{equation}\label{eqn:kick_bends}
\Delta x', \Delta y'\approx \q{0.5}{nrad}.
\end{equation}

%\begin{table*}
%	\caption{\label{tab:md_param}%
%		Beam parameters and machine configuration as used in the two resonant excitation experiments in 2016 and 2017 \cite{resexmd2016,resexmd2017}. The plane of the excitation is abbreviated with H for horizontal, V for vertical and H+V horizontal and vertical at the same time.
%	}
%	\begin{ruledtabular}
%		\begin{tabular}{lcc}
%			Parameter & Experiment 2017 & Experiment 2016  \\
%			\colrule
%			beam &\multicolumn{2}{c}{Beam~1} \\
%			beam energy &\multicolumn{2}{c}{injection energy, 450 GeV} \\\hline
%			single bunch intensity &\multicolumn{2}{c}{$0.7\times10^{11}$} \\
%			normalized emittance &\multicolumn{2}{c}{$2.5-3.5~\mathrm{\mu m}$} \\
%			$4\sigma$ bunch length & \multicolumn{2}{c}{1.3~ns}\\
%			$1\sigma$ bunch length & \multicolumn{2}{c}{9.7~cm}\\
%			number of bunches & $3\times72=216$ bunches & $12\times4=48$ single bunches\\
%			& (+ 1 pilot + 12 nominal) & \\\hline
%			injection optics, $\beta^*=11$~m & standard optics 2017 & standard optics 2016\\
%			Landau damping octupoles  & \multicolumn{2}{c}{$I_{\mathrm{MO}}=\pm19.6~\mathrm{A}$, explicitly +19.6 A for MOF circuit and}\\
%			& \multicolumn{2}{c}{-19.6~A for MOD circuit (standard 2016 settings)}\\\hline
%			working point $(Q_x,Q_y)$ & (62.27,60.295) & (64.28,59.31)\\
%			chromaticity $(Q'_x,Q'_y)$ & \multicolumn{2}{c}{(+15,+15)}\\\hline
%			pulsing patterns  & $7^{\mathrm{th}}$~turn H,V,H+V;& $7^{\mathrm{th}}$~turn H;\\
%			& $8^{\mathrm{th}}$~turn H,V,H+V;& $10^{\mathrm{th}}$~turn V;\\
%			& random  H,V,H+V;& \\
%		\end{tabular}
%	\end{ruledtabular}
%\end{table*}

@@

\section{Experimental and simulation setup\label{sec:exp}}
%put picture on top
\begin{figure*}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		2016 experiment\\
		\includegraphics[width=0.9\linewidth]{bunchfilling_2016.png}	
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		$ $\\
		2017 experiment\\
		\includegraphics[width=0.9\linewidth]{bunchfilling_2017.png}	
	\end{minipage}
	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom). In 2016 a total of 48~bunches was used and in 2017 a total of 216~bunches. Each bunch is indicated with a cyan vertical bar. The bunches were grouped in sub-groups of 4~for 2016 and 6~for 2017 bunches, where each sub-group experienced the same excitation pattern and amplitude. The excitation amplitude is shown in blue and red. In 2016 the excitation was only applied in one plane. In 2017 more bunches could be injected without compromising machine protection making it possible to test all three planes in one fill.}
\end{figure*}
%put picture on top

>>>>

Because no HEL is currently installed in the LHC, the kicks on the
proton beam core must be emulated by other devices to determine their
effects experimentally in a given machine and to guide design and
tolerances. In particular, during the experiments presented in this
paper, the LHC transverse damper system (described in detail in
Section~\ref{sec:adt}) was used for this purpose. This system can
generate transverse dipole kicks with a wide range of excitation
patterns.

For comparison with experiments and simulations, here we estimate to
first order the magnitude of the dipole kicks that may be expected
from the HEL. As one can see below, the contribution from the central
region is in general dominating.

\subsection{Uncompensated kicks from injection and extraction bends}
\label{core:sec:1}

To estimate the dipole component of the kicks that originate from the
injection and extraction bends, we used the approach described in
Ref.~\cite{hel_bends_stancari}. The bends are modeled as a bent
cylindrical pipe with a static charge distribution of electrons. The
resulting electric field is calculated, with the vacuum pipe as
boundary, using the solvers of the Warp particle-in-cell
code~\cite{warp}. (The contribution of the magnetic field generated by
the electron current was neglected in this study.) The field,
integrated over the trajectory of the protons, is then translated into
a symplectic kick map.

In case of a U-shaped electron lens, where electron gun and collector
are on the same side, the transverse (and possibly pulsed) dipole
kicks generated by the electron charge at the entrance and exit add
up. For an S-shaped electron lens, on the other hand, where gun and
collector are on opposite sides, these kicks compensate each
other. For this reason, an S-shape was chosen for the HL-LHC HEL
(Fig.~\ref{fig:hel_layout}). A disadvantage of the S-shape is that the
static magnetic kicks due to the toroidal sections do add up, but they
can be compensated by conventional dipole correctors, especially in a
high-energy machine.

In the case under study of an S-shaped electron lens, residual
uncompensated kicks arise from differences in electron charge
distribution between the injection and extraction bends. Here we
conservatively assume 10\% fluctuations between the entrance and exit
and, furthermore, that these differences add up.

The maximum values of the integrated electric fields calculated in
Ref.~\cite{hel_bends_stancari}, based upon an electron beam of 1~A at
5~keV, are
%
\begin{equation}
  \int_{z_1}^{z_2} E_{x,y} \, dz= \q{10}{kV}.
\end{equation}
%
Scaling to HL-LHC and HEL design parameters
(Table~\ref{tab:hllhc_param} and Table~\ref{tab:hel_param}) yields the
following integrated field and corresponding kick:
%
\begin{equation}
  \int_{z_1}^{z_2} E_{y} \, dz = \q{36}{kV} \Rightarrow \Delta y' = \q{5.1}{nrad},
\end{equation}
%
as described in detail in
Ref.~\cite{md_sim_hel_res_ex_fitterer}. Assuming a residual difference
of 10\% between entrance and exit, the expected kick is approximately
%
\begin{equation}
  \label{eqn:kick_bends}
  \Delta x', \Delta y' = \q{0.5}{nrad}.
\end{equation}

@@

\section{Experimental and simulation setup\label{sec:exp}}
% put picture on top

\begin{figure*}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		2016 experiment\\
		\includegraphics[width=0.9\linewidth]{bunchfilling_2016.png}	
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		$ $\\
		2017 experiment\\
		\includegraphics[width=0.9\linewidth]{bunchfilling_2017.png}	
	\end{minipage}
	\caption{\label{fig:fill} Filling and excitation scheme for LHC experiments in 2016 (top) and 2017 (bottom). In 2016 a total of 48~bunches was used and in 2017 a total of 216~bunches. Each bunch is indicated with a cyan vertical bar. The bunches were grouped in sub-groups of 4~for 2016 and 6~for 2017 bunches, where each sub-group experienced the same excitation pattern and amplitude. The excitation amplitude is shown in blue and red. In 2016 the excitation was only applied in one plane. In 2017 more bunches could be injected without compromising machine protection making it possible to test all three planes in one fill.}
\end{figure*}
%put picture on top
